---
title: "Predicting TB Medication Adherence Using a Risk Score Model"
date: "February 2024"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
#devtools::install_github("hjeglinton/riskscores", build_vignettes = TRUE)

library(knitr)
library(xtable)
library(glmnet)
library(pROC)
library(tableone)
library(glmnet)
library(caret)
library(gridExtra)
#library(kableExtra)
library(gtsummary)
library(riskscores)
#source('risk.R')

set.seed(1)

```

# Data Preprocessing 



```{r}

raw_data <- read.csv("../../data/tb/Peru_TB_data.csv") 

```





```{r}
setwd("../../R")
source('tb_preprocessing.R')
tb_df <- tb_preprocessing(raw_data) 
tb_mat <- tb_as_matrix(tb_df)

```

# Model Fitting 

```{r, out.width="75%", fig.align = 'center', warning = FALSE}

X <- as.matrix(tb_mat[,-ncol(tb_mat)])
y <- tb_mat[,ncol(tb_mat)]

# CV
# get folds
folds <- stratify_folds(y, nfolds = 5, seed = 5)


lasso_cv <- cv.glmnet(x=X[,-1], y=y, alpha=1, type.measure = "deviance")

cv_results <- cv_risk_mod(X, y, foldids = folds, a = -10, b = 10, nlambda = 25)

plot(cv_results, lambda_text = FALSE) + 
  labs(title = "Figure 1. Cross Validation Results")


```



```{r, warning = FALSE}

mod_cd <- risk_mod(X, y, lambda0 = exp(-5.7), a = -10, b = 10)
mod_l0 <- risk_mod(X, y, lambda0 = exp(-6.7), a = -10, b = -10)

```







# Model Evaluation 


We compared the performance of our risk score model to logistic regression, lasso regression, and rounded logistic regression. The coefficients for each of these models are reported in Table 3. As expected, the logistic regression model assigns a non-zero coefficient value to each covariate, while the lasso model shrinks many coefficients to zero. 

```{r, include = FALSE}

coef_vals <- matrix(0, ncol=4, nrow=ncol(X)-1)

# risk model prediction
coef_cd <- coef(mod_cd)
cd_probs <- predict(mod_cd, type = "response")


# lasso 
lasso_res <- cv.glmnet(x=X[,-1], y=y, alpha=1)
lasso_mod <- glmnet(x=X[,-1], y=y, lambda=lasso_res$lambda.min, alpha=1, foldids = folds)
coef_lasso <- coef(lasso_mod)
lasso_probs <- as.vector(predict(lasso_mod, newx=X[,-1]))

# Rounded Lasso 
scalar <- ifelse(is_empty(coef_lasso[coef_lasso != 0][-1]),
                 1,
                 median(coef_lasso[coef_lasso != 0][-1], na.rm = TRUE))
b0_lasso_scaled <- coef_lasso[1]/scalar
coef_lasso_rounded <- c(b0_lasso_scaled, 
                        round(coef_lasso[-1]/scalar, 0))
scores <- as.vector(X %*% coef_lasso_rounded)
mod_scores <- glm(y ~ scores, family = "binomial")
lasso_rounded_probs <- predict(mod_scores, type = "response")


# discrimination
cd_roc <- roc(factor(tb_df$adherence_outcome), cd_probs[,1], quiet = TRUE) %>% ci.auc()
lasso_roc <- roc(factor(tb_df$adherence_outcome), lasso_probs, quiet = TRUE) %>% ci.auc()
round_roc <- roc(factor(tb_df$adherence_outcome), lasso_rounded_probs, quiet = TRUE) %>% ci.auc()


```






```{r}
data.frame(l_ci = c(lasso_roc[[1]], 
                   round_roc[[1]], 
                   cd_roc[[1]]),
           auc = c(lasso_roc[[2]], 
                   round_roc[[2]], 
                   cd_roc[[2]]),
           u_ci = c(lasso_roc[[3]], 
                   round_roc[[3]], 
                   cd_roc[[3]]),
           row.names = c("Lasso", "Rounded Lasso", "RiskCD")) %>%
  round(4)
```


```{r}
data.frame(lasso = as.vector(coef_lasso),
           rounded_lasso = as.vector(coef_lasso_rounded),
           cd = as.vector(coef_cd),
           row.names = dimnames(X)[[2]]) %>%
  slice(-1) %>%
  filter_all(any_vars(. != 0)) %>%
  kable(digits = 3, col.names = c("Lasso", "Rounded Lasso", "CD"),
      booktabs = T, caption = "Model Coefficients") 
```




```{r, out.width = "75%", fig.align='center'}


predict_df <- data.frame(score = predict(mod, type = "score"),
                         response = predict(mod, type = "response"))

score_v_risk <- ggplot() + 
  #geom_point(data = predict_df, aes(x = score, y = response), size = 1) +
  geom_function(data = data.frame(x = seq(-8,12)), aes(x), 
                fun = function(x) get_risk(mod, x), size = 1, alpha = 0.9) + 
  labs(x = "RiskCD Score", y = "Risk of Suboptimal Adherence", title = "") +
  theme_bw() +  
  theme(axis.text = element_text(size = 10), panel.grid.minor.y = element_blank()) + 
  scale_x_continuous(breaks = seq(-8, 12, 2)) + 
  scale_y_continuous(breaks = seq(0, 1, 0.1))

ggsave(score_v_risk, width = 7, height = 5, dpi = 300, filename = "score_v_risk_adherence.png")

```


# Conclusion

Despite a relatively small sample size, our example model output made sense intuitively. Our model identified higher stigma and having to take the medication in-person at a clinic as factors increasing the risk of non-adherence. Having never had COVID and having family supervision of treatment decreased the risk of non-adherence. Although the performance of this model should be validated with new data, it may still help with identifying which factors are the most important in predicting non-adherence. 


\pagebreak

# EDA Appendix



### Figure A1. Outcome Distributions
```{r, warning = FALSE, message = FALSE, fig.height = 3}
# outcome distributions
p1 <- ggplot(raw_data) +
  geom_histogram(aes(x = PCTadherence)) + 
  lims(y = c(0, 210)) + 
  theme_bw()

p2 <- ggplot(raw_data) +
  geom_histogram(aes(x = PCTadherence_sensi)) + 
  lims(y = c(0, 210)) + 
  theme_bw()

grid.arrange(p1, p2, ncol = 2)
```

### Figure A2. Continuous Covariate Distributions

```{r, warning = FALSE, message = FALSE, fig.height = 5}
p3 <- ggplot(raw_data) +
  geom_histogram(aes(x = self_eff)) + 
  theme_bw()

p4 <- ggplot(raw_data) +
  geom_histogram(aes(x = tx_mos)) + 
  theme_bw()

p5 <- ggplot(raw_data) +
  geom_histogram(aes(x = audit_tot)) + 
  theme_bw()

p6 <- ggplot(raw_data) +
  geom_histogram(aes(x = stig_tot)) + 
  theme_bw()

p7 <- ggplot(raw_data) +
  geom_histogram(aes(x = phq9_tot)) + 
  theme_bw()

p8 <- ggplot(raw_data) +
  geom_histogram(aes(x = age_BLchart)) + 
  theme_bw()

p9 <- ggplot(raw_data) +
  geom_histogram(aes(x = ace_score)) + 
  theme_bw()

grid.arrange(p3, p4, p5, p6, p7, p8, ncol = 2)
```

### Table A1. Processed Variable Summary

```{r, message = FALSE}
tbl_summary(tb_df)#%>%
   #as_kable_extra(booktabs = TRUE, 
                # longtable = TRUE) #%>%
 # kableExtra::kable_styling(font_size = 9, 
                           # latex_options = c("repeat_header", "HOLD_position"))

```


```{r, eval = FALSE, message = FALSE}


# load risk model files
X <- tb_matrix[, -ncol(tb_matrix)]
y <- tb_matrix[, ncol(tb_matrix)]

coef_vals <- matrix(0, ncol=3, nrow=ncol(X)-1)

# risk model prediction 
risk_output_cv <- cv_risk_mod(X, y, a=-5, b=5)
risk_output <- risk_mod(X, y, a=-5, b=5, lambda0=risk_output_cv$lambda_min)
coef_vals[,1] <- risk_output$beta[-1]
risk_probs <- predict(risk_output$glm_mod, type="response")
risk_pred <- as.factor(ifelse(risk_probs < 0.5, FALSE, TRUE))

# glm prediction
glm_mod <- glm(y~X-1, family = "binomial")
coef_vals[, 2] <- coef(glm_mod)[-1]
glm_probs<- predict(glm_mod, type="response")
glm_pred <- as.factor(ifelse(glm_probs < 0.5, FALSE, TRUE))

# lasso prediction

lasso_res <- cv.glmnet(x=X[,-1], y=y, alpha=1)
lasso_mod <- glmnet(x=X[,-1], y=y, lambda=lasso_res$lambda.min, alpha=1)
coef_vals[,3] <- coef(lasso_mod)[-1]
lasso_probs <- as.vector(predict(lasso_mod, newx=X[,-1]))
lasso_pred <- as.factor(ifelse(lasso_probs < 0.5, FALSE, TRUE))

print(coef_vals)

# confusion matrices
confusionMatrix(as.factor(tb_df$adherence_outcome), risk_pred)
confusionMatrix(as.factor(tb_df$adherence_outcome), glm_pred)
confusionMatrix(as.factor(tb_df$adherence_outcome), lasso_pred)

# discrimination
roc(as.factor(tb_df$adherence_outcome), risk_probs)
roc(as.factor(tb_df$adherence_outcome), glm_probs)
roc(as.factor(tb_df$adherence_outcome), lasso_probs)

# find risk score probs to summarize
tb_df$scores <- (X[,-1] %*% risk_output$beta[-1])
ggplot(tb_df)+geom_histogram(aes(x=scores, fill=adherence_outcome), alpha=0.5)

# for each score find predicted probability and also find percent class 1
range_scores <- range(tb_df$scores)
all_scores <- seq(range_scores[1], range_scores[2])
vals <- risk_output$gamma*(risk_output$beta[1]+range_scores)
probs <- exp(vals)/(1+exp(vals))
props <- tb_df %>% 
  mutate(rnd_scores = floor(scores)) %>%
  group_by(rnd_scores) %>% 
  summarize(prop = sum(adherence_outcome)/n()) 

ggplot()+geom_line(aes(x=range_scores, y=probs)) + 
  geom_point(aes(x=props$rnd_scores,y=props$prop)) + 
  labs(x="Score", y="Predicted or Observed Probability")



```


\pagebreak

# Code Appendix



```{r  ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE, include=TRUE}

```


